---
title: "Data Manipulation with dplyr"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "February 28, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr);
df <- read_tsv (file = "/home/utilisateur/proglitt/TD5/tmp/dpt2015.txt",
                locale = locale(encoding = "ISO-8859-1"));
```

```{r}
library(xtable);
df %>% sample_n(6) %>% as.data.frame() %>% xtable(.);

df;
```



```{r}
df %>% head(n=2);
```

```{r}
df %>% select(preusuel);
```
```{r}
df %>% filter(annais == 1948);
```

```{r}
df %>% arrange(annais);
```

```{r}
df %>% filter(preusuel == 'HANENE');
```


```{r, fig.width=10, fig.height=5}
library(ggplot2);
df %>%
   group_by(annais) %>%
   summarize(somme=sum(preusuel == "LAURENT")) %>%
   filter(annais>1980) %>%
   ggplot(aes(x=annais, y=somme)) +
          geom_point() + ylim(0,NA) + theme_bw();
```

Avec Hanene HEDRACHE, on a d'abord représenté le nombre de personnes s'appelant Laurent qui sont nés sur la période de 1900 à 2015. Pour des raisons de lisibilité, nous avons réduit cette période de 1980 à 2015. Cela traduit ainsi une décroissance des Laurent.



```{r, fig.width=10, fig.height=5}
library(ggplot2);
df %>%
   group_by(dpt) %>%
   filter(dpt>80,dpt<99) %>%
   summarize(nb=sum(nombre)) %>%
   ggplot(aes(x=dpt, y=nb)) +
          geom_point() + ylim(0,NA) + theme_bw();
```

Ici nous avons représenté les nombres de naissances par département. Ils sont en général constants mais des départements se démarquent (le 13, le 59, le 69, le 75, etc., le 99 ne comptant pas et le 97 étant les dom tom)




This is a demonstration of how dplyr works.

First, we need some data.

```{r}
df <- data.frame(
  name = c("John", "Mary", "Alice", "Peter", "Roger", "Phyllis"),
  age = c(13, 15, 14, 13, 14, 13),
  sex = c("Male", "Female", "Female", "Male", "Male", "Female")
)
head(df);
```

Load the necessary packages:

```{r}
library(dplyr);
library(magrittr);
```

Learning how to use the pipe operator (%>%):

```{r}
df %>% head(n=2);
```

You can chain multiple pipes:

```{r}
df %>% head(n=2) %>% nrow;
```

Now, let's use dplyr verbs, starting with _select()_ to select only one column:

```{r}
df %>% select(name);
```

You can also remove some column using the minus operator:

```{r}
df %>% select(-sex);
```

Now, let's use the _filter()_ verb (let's say we need only those of age 13):

```{r}
df %>% filter(age == 13);
```

If we need to re-order, we can use _arrange()_ (you can pass multiple variables as well):

```{r}
df %>% arrange(age);
```

And to create new columns, we can _mutate()_ (you can use existing variable names):

```{r}
df %>% mutate(birth = 2017 - age);
```

The _summarize()_ verb can be used to reduce data (to know the average age, for instance):

```{r}
df %>% summarize(mean_age = mean(age));
```

The _group_by()_ verb can be used to split data, apply some function and combine results (for example, to know the average age depending on sex):

```{r}
df %>% group_by(sex) %>% summarize(mean_age = mean(age));
```

We can also _group_by()_ by multiple variables (to see the number of occurrences of age/sex combinations):

```{r}
df %>% group_by(age, sex) %>% summarize(N=n());
```

How to use in combination with _ggplot()_:

```{r, fig.width=2.5, fig.height=2}
library(ggplot2);
df %>%
   group_by(sex) %>%
   summarize(mean_age = mean(age)) %>%
   ggplot(aes(x=sex, y=mean_age)) +
          geom_point() + ylim(0,NA) + theme_bw();
```
































